---
title: "Data Cleaning"
author: "Karine Almeida"
date: "2023-07-19"
output: pdf_document
---

# Business Understanding  
  
A fictitious car buying and selling company is having difficulties reselling 
used cars in its catalogue. With the aim of pricing your catalog more competi-
tively and thus recovering the poor performance in this sector, I will analyze
the data to answer business questions and create a predictive model that prices 
the customer's cars so that they are as close to market values. 
**In this notebook you will have access to a descriptive analysis of the data,**
**insights and answers to some business questions.**    
   
# Data cleaning 
  
## Installing packages   
```{r message=FALSE, warning=FALSE}
pacotes <- c('tidyverse','knitr','kableExtra', 'ggplot2', "paletteer",
             "scales","DT", "kableExtra", 'gridExtra', 'xlsx')

options(rgl.debug = TRUE)

if(sum(as.numeric(!pacotes %in% installed.packages())) != 0){
  instalador <- pacotes[!pacotes %in% installed.packages()]
  for(i in 1:length(instalador)) {
    install.packages(instalador, dependencies = T)
    break()}
  sapply(pacotes, require, character = T) 
} else {
  sapply(pacotes, require, character = T) 
}

```
  
## Viewing the database  
```{r echo=TRUE, warning=FALSE}
cars_train <- read_delim("../data/cars_train.txt", show_col_types = FALSE)
glimpse(cars_train)
```

## Identifying and removing missing data  
```{r}
sapply(cars_train, function(x) sum(is.na(x)))

```
## veiculo_único_dono: I will consider that the missing values represent "mais de um dono" (more than one owner), so I will replace it with that.
```{r}
cars_train$veiculo_único_dono <- 
  cars_train$veiculo_único_dono %>% replace_na("mais de um dono")
```

## veiculo_licenciado: Null values were considered as "não licenciado"
```{r}
cars_train$veiculo_licenciado <- 
  cars_train$veiculo_licenciado %>% replace_na("não licenciado")
```

## dono_aceita_troca: missing values replaced by "não aceita troca"
```{r}
cars_train$dono_aceita_troca <- 
  cars_train$dono_aceita_troca %>% replace_na("não aceita troca")
```

## ipva_pago: missing values replaced by "IPVA não pago"
```{r}
cars_train$ipva_pago <- 
  cars_train$ipva_pago %>% replace_na("ipva não pago")
```

```{r}
cars_train$num_fotos <- 
  cars_train$num_fotos %>% replace_na(0)
```


The missing items identified in the other variables, in addition to being very 
numerous, are redundant because even if corrected they would provide information
without variation. These variables will be deleted.

## Removing variables with redundant values, excess missing values and many categories
```{r}
cars_train <- subset(cars_train, 
                     select = -c(id,veiculo_alienado,revisoes_concessionaria,
                     garantia_de_fábrica,revisoes_dentro_agenda,
                     elegivel_revisao))
```

## Check whether the levels of the categorical variables are balanced
```{r}
#The brands FERRARI, IVECO, JAC, BRM and EFFA have very few observations, which therefore does not have enough variability to contribute to the model. Therefore, they will be excluded.

#marca
cars_train <- subset(cars_train, marca != "FERRARI")
cars_train <- subset(cars_train, marca != "IVECO")
cars_train <- subset(cars_train, marca != "JAC")
cars_train <- subset(cars_train, marca != "BRM")
cars_train <- subset(cars_train, marca != "EFFA")

#ano_de_fabricação
cars_train <- subset(cars_train, ano_de_fabricacao != "1985")
cars_train <- subset(cars_train, ano_de_fabricacao != "1988")
cars_train <- subset(cars_train, ano_de_fabricacao != "1990")

#ano_modelo
cars_train <- subset(cars_train, ano_modelo != "1997")
cars_train <- subset(cars_train, ano_modelo != "2006")
cars_train <- subset(cars_train, ano_modelo != "2008")
cars_train <- subset(cars_train, ano_modelo != "2010")

#num_portas
cars_train <- subset(cars_train, num_portas != 3)

#cor
cars_train <- subset(cars_train, cor != "Dourado")
cars_train <- subset(cars_train, cor != "Verde")
cars_train <- subset(cars_train, cor != "Vermelho")

#estado_vendedor
cars_train <- subset(cars_train, estado_vendedor != "Roraima (RR)")
cars_train <- subset(cars_train, estado_vendedor != "Maranhão (MA)")
cars_train <- subset(cars_train, estado_vendedor != "Rondônia (RO)")
cars_train <- subset(cars_train, estado_vendedor != "Piauí (PI)")

#anunciante
cars_train$anunciante[cars_train$anunciante == "Acessórios e serviços para autos"] <- "Concessionária"
cars_train$anunciante[cars_train$anunciante == "Loja"] <- "Concessionária"


#num_portas, blindado, entrega_delivery, troca, dono_aceita_troca: These are variables that present a suspicious distribution of the data. Before doing the modeling I will investigate which of them should or should not be excluded.
pie(table(cars_train$num_portas), main = "tipo", col= c("#003154","#dd4a37"))
pie(table(cars_train$blindado), main = "tipo", col= c("#003154","#dd4a37"))
pie(table(cars_train$entrega_delivery), main = "tipo", col= c("#003154","#dd4a37"))
pie(table(cars_train$troca), main = "tipo", col= c("#003154","#dd4a37"))
pie(table(cars_train$dono_aceita_troca), main = "tipo", col= c("#003154","#dd4a37"))

```

## Identifying outliers
```{r}
boxplot(cars_train$preco, las=2, xlab="preco") 
boxplot(cars_train$hodometro,las=2, xlab="hodometro")
```
 
## Quartile function
```{r}
quartil <- function(column){
  
  q1 <- quantile(column, 0.25, na.rm = TRUE) #1º quartil
  q3 <- quantile(column, 0.75, na.rm = TRUE) #3º quartil
  iq <- q3 - q1 #interquartil
  lim_sup <- q3 + 1.5*iq #limite superior
  return(lim_sup)
}
```

## Calculating outliers across the top quartile  
```{r}
max_preco <- quartil(cars_train$preco)
max_hodo <- quartil(cars_train$hodometro)
print(paste("Preço:",max_preco, "Hodometro:", max_hodo))
```
## Now we will discard the lines where price and odometer are above the upper limit 
```{r}
for (i in seq_along(cars_train$preco)){
  if (cars_train$preco[i] > 295085.69){
    cars_train$preco[i] <- mean(cars_train$preco)
  } 
}

for (i in seq_along(cars_train$hodometro)){
  if (cars_train$hodometro[i] > 158264){
    cars_train$hodometro[i] <- mean(cars_train$hodometro)
  } 
}
```
  
## Visualize the distribution of variables, now without the outliers
```{r}
boxplot(cars_train$preco, las=2, xlab="preco") 
boxplot(cars_train$hodometro,las=2, xlab="hodometro")
```

## Transforming variables into factors
```{r}
cars_train[,c("marca", "cor", "cambio", "tipo", "ano_de_fabricacao",
              "ano_modelo", "dono_aceita_troca", "veiculo_único_dono",
              "ipva_pago", "veiculo_licenciado","blindado",
              "tipo_vendedor","estado_vendedor", 
              "anunciante")]= 
  lapply(cars_train[,c("marca", "cor", "cambio", "tipo", "ano_de_fabricacao",
              "ano_modelo", "dono_aceita_troca", "veiculo_único_dono",
              "ipva_pago", "veiculo_licenciado","blindado",
              "tipo_vendedor","estado_vendedor", 
              "anunciante")], as.factor)

summary(cars_train)
```
## Saving clean dataset
```{r}
write.csv(cars_train, "../data/cars_train_clean1.csv", row.names = FALSE)
```



